\section{Introduction}

Much operational computer work is \emph{repeated}: the business intent is stable,
the sequence is mostly known, errors are costly, and the same task may run
hundreds of times. Reasoning through such a workflow on every run is both
wasteful and unsafe. Wasteful, because a general model re-derives the same plan
each time, paying latency, inference cost, and run-to-run variation. Unsafe,
because the usual success signal is the \emph{screen}: a model or a script sees a
rendered ``Saved'' banner and stops---but a banner is not a persisted write. The
same green pixels can follow a partial save, a duplicate submission, a stale
overwrite, or a silently rejected transaction. The failure mode that matters in
production is not the visible crash; it is the \emph{silent wrong action} that
reports success.

Graphical user interfaces remain the practical integration boundary for legacy,
third-party, and remote applications, so this problem cannot be assumed away with
an API. Selector-based automation is fast and deterministic when stable structure
is available but brittle under interface drift, while visual scripting extends to
pixel surfaces \cite{sikuli}. Recent computer-use agents instead observe the
screen and choose each action with a general model
\cite{anthropic2024computeruse,uitars2025,cheng2024seeclick}, a useful
formulation for previously unseen tasks and broad benchmarks
\cite{webarena,workarena,osworld}. Neither line closes the safety gap: RPA does
not check the system of record, and an agent re-reasons every run yet still trusts
the screen.

\system{} treats a successful human demonstration as source material for a
\emph{program} rather than as context for a model on every run
(Figure~\ref{fig:pipeline}). Compilation extracts parameters, redundant target
evidence, postconditions, and risk metadata. Replay follows the compiled program
deterministically and makes zero model calls on the healthy path; models are
optional repair tiers, not the controller. Where a step is consequential, an
independent verifier checks the effect against the application's own
system of record, and the runtime refuses rather than guessing when it cannot
verify. The distinction the paper foregrounds is not whether a task
\emph{completed} but whether identity and external effects were \emph{verified},
and whether the system can \emph{refuse} when evidence is insufficient. In an
injected-fault study this distinction is decisive, and we measure it end to end
through the live replayer and judge it by an independent system-of-record ground
truth: screen-only verification silently accepted 54 of 90 fault runs, an
out-of-band effect check cut that to 9 of 90, and a complete system-of-record
read path to 0 of 90 (Section~\ref{sec:results}). The same gap replicates on a
second, non-healthcare lending domain.

This paper makes five contributions:
\begin{enumerate}
  \item A demonstration compiler and backend-neutral workflow representation
        for deterministic repeated GUI execution.
  \item A resolution ladder that prefers structured identity, then local visual
        evidence, and records bounded drift repairs as reviewable changes.
  \item System-of-record effect verification that treats the screen as a weak
        oracle and the application's own state as the strong oracle, closing the
        gap between a rendered success and a persisted effect.
  \item Governance mechanisms that distinguish runnable from certified, verify
        identity and system effects where configured, and refuse rather than
        silently degrade.
  \item An evaluation protocol that reports task success \emph{together with}
        silent incorrect success and over-halt, with a failure taxonomy and
        \emph{machine-checked paper constants} that bind every headline number
        to a released raw artifact or bounded aggregate summary. Raw grown
        corpora, deployment-derived tuning, and target-specific recipes remain
        outside the public release.
\end{enumerate}

The evaluation deliberately separates a broad product architecture from the
scope of each measured result. Browser workflows supply the comparative and
breadth studies. Additional task-bound qualifications cover Windows UIA,
native macOS, and real-network RDP with independent effect oracles and ambiguity
or stale-target refusals. These results establish working substrate mechanisms
for the named environments; they are not a population estimate for arbitrary
enterprise applications. Citrix ICA/HDX is a pixel-only remote-display
environment in the same class as the measured RDP backend, but no Citrix result
is reported here: the qualification covers RDP over Aardwolf into a Parallels
Windows guest, not an ICA/HDX session, and RDP evidence does not transfer to
Citrix. The limitations section states this scope precisely and names the
specific validation it would take.
