\section{Related Work}

\paragraph{GUI automation and RPA.}
Robotic process automation systems execute rules over existing application
interfaces and organizational workflows; their adoption and governance
questions are surveyed by van der Aalst et al. \cite{vanderaalst2018rpa}.
Screenshot-driven automation predates current multimodal agents. Sikuli made
visual patterns a first-class interface for search and scripting
\cite{sikuli}. \system{} shares the objective of operating interfaces that lack
a practical API, but combines structured elements and visual evidence in an
ordered ladder and makes a repair an auditable program change.

\paragraph{Programming by demonstration and synthesis.}
Programming by demonstration infers reusable behavior from user actions rather
than requiring conventional source code \cite{cypher1993watch}. Programming-by-
example systems such as FlashFill demonstrate how compact examples can produce
deterministic programs while exposing ambiguity in the inferred intent
\cite{gulwani2011automating}, and web-scraping-by-demonstration systems such as
Rousillon compile direct manipulation into a reusable program
\cite{chasins2018rousillon}. A GUI trace similarly under-specifies intent:
coordinates do not identify the semantic target, and one trajectory does not
determine every branch or loop. \system{} therefore retains redundant target
evidence, represents control flow explicitly, and quarantines underdetermined
induction instead of filling gaps with an unverified guess. It also departs from
this line in two ways: it repairs targets at runtime under drift, and it verifies
downstream effects rather than only reproducing input actions.

\paragraph{Computer-use agents and benchmarks.}
A rapidly growing line of computer-use agents observes the screen and chooses
each action with a general multimodal model
\cite{anthropic2024computeruse,uitars2025,cheng2024seeclick}, evaluated on
breadth-oriented benchmarks such as WebArena, WorkArena, and OSWorld
\cite{webarena,workarena,osworld}. These emphasize task completion under online
reasoning on previously unseen tasks. Our question is complementary: after a
workflow has been \emph{demonstrated}, when can execution be compiled so that
repeated healthy runs need no model decision? We retain task-level oracles but
additionally measure silent incorrect success and over-halt, which distinguish
unsafe success from conservative refusal---quantities these benchmarks do not
report.

\paragraph{Runtime verification, oracles, and end-to-end effects.}
Runtime verification synthesizes monitors that evaluate execution traces
against explicit safety properties \cite{havelund2002synthesizing}, and the
end-to-end argument places a correctness check at the layer that can establish
the intended property rather than an intermediate subsystem \cite{saltzer1984end}.
Deciding whether an observed outcome is correct is the \emph{test-oracle problem}
\cite{barr2015oracle}; the faults \system{} targets---duplicate submission,
double delivery, stale overwrite---are the same exactly-once and idempotency
hazards studied in transactional systems \cite{gray1981transaction}. These
principles motivate \system{}'s separation between screen postconditions and
system-of-record effects: a rendered success banner cannot independently
establish that a transaction was persisted correctly. Unlike a general
formal-verification claim, the guarantees here remain scoped to armed steps,
declared effects, configured verifiers, and named policies.

\begin{table}[t]
\centering
\caption{Where \system{} sits. Columns are the properties that matter for
repeated, consequential GUI work; \checkmark{} = designed-in, $\sim$ = partial
or tool-dependent, blank = not addressed by the paradigm.}
\label{tab:positioning}
\small
\begin{tabular}{lccccc}
\toprule
 & No per-run & Deterministic & Repairs & Verifies & Refuses on \\
Approach & model call & healthy path & under drift & effect & ambiguity \\
\midrule
Selector RPA \cite{vanderaalst2018rpa} & \checkmark & \checkmark & & & \\
Visual scripting \cite{sikuli}          & \checkmark & \checkmark & $\sim$ & & \\
PBD / synthesis \cite{cypher1993watch,gulwani2011automating} & \checkmark & \checkmark & & & $\sim$ \\
Computer-use agents \cite{anthropic2024computeruse,webarena} & & & \checkmark & & $\sim$ \\
\textbf{\system{}} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table}
