\section{System Design}

\subsection{Record and compile}
A recorder captures actions, frames, element observations, and structural state
when available. Declared secret fields are replaced with runtime parameters.
The compiler emits a bundle containing a workflow program, target crops,
locators, landmarks, postconditions, parameter declarations, and a manifest.
The linear trace is the simplest program; the intermediate representation also
supports states, guarded transitions, branches, loops, and subflows. Induction
that cannot determine a branch condition quarantines the result rather than
inventing one.

\subsection{Resolution ladder}
For an action target, replay tries evidence in descending order of fidelity:
structured DOM or accessibility identity, local and global template matching,
OCR, landmark geometry, and an optional grounding model. A backend presents a
small common contract for observation and actuation. This provides a pixel
floor without discarding structured evidence on browsers and native desktops.

Healthy execution stops at deterministic tiers and makes zero model calls. A
lower rung resolving a changed but verified target creates a repair patch. The
runtime can save a healed derivative bundle, but does not silently replace the
deployed source bundle.

\subsection{Postconditions and reports}
Each step may carry visual or structural postconditions such as text presence,
region stability, URL change, or window state. Replay writes a human-readable
and machine-readable report with rung, confidence, identity verdict,
postconditions, effects, model calls, elapsed time, repair, and halt reason.
Postconditions are necessary but not sufficient: a success banner is evidence
about the screen, not independent evidence about a database write.

\subsection{Backends}
The browser backend is the reference implementation and supplies published
end-to-end evidence against a real third-party application. Windows UI
Automation adds structured candidate enumeration and native actions behind a
typed, authenticated in-session contract. Native macOS binds capture and input
to a unique application window. The RDP backend exposes the same runtime
contract over a pixel-only network session. Section~4 reports scoped task and
effect qualifications for all three mechanisms. Citrix/ICA/HDX remains a
separate deployment environment and does not inherit RDP qualification.
