\section{Governed Execution}

\subsection{Identity before action}
Target resolution answers where to act; identity verification asks whether the
resolved target belongs to the intended entity. Structured text is authoritative
where available. On pure-pixel substrates, OCR ambiguity can make distinct
identifiers observationally identical. The safe response is abstention and
halt, not a guessed click. Importantly, identity guarantees apply only to armed
steps. Current real bundles can contain unarmed clicks, so coverage is reported
per workflow and certification policy can impose a minimum.

\subsection{Effect verification}
A screen may paint success after a partial, duplicate, rejected, or stale write.
Consequential steps can therefore declare typed effects and use an independent
verifier against REST, FHIR, or document state. The verifier snapshots the
system of record before and after an action and returns confirmed, refuted, or
indeterminate. A declared effect without a configured verifier halts. Effects
are currently authored per deployment; the compiler does not infer them from a
single demonstration. This is the classic \emph{oracle problem}
\cite{barr2015oracle} made concrete: the screen is a weak oracle and the
application's own state is the strong oracle (Figure~\ref{fig:oracle}). Many of
the faults it catches---duplicate submission, double delivery, stale
overwrite---are exactly-once and idempotency hazards long studied in
transactional systems \cite{gray1981transaction}.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  font=\small,
  screen/.style={draw=oaink, rounded corners=2pt, minimum width=30mm,
    minimum height=13mm, align=center, fill=white},
  db/.style={draw=oaink, cylinder, shape border rotate=90, aspect=0.28,
    minimum width=22mm, minimum height=15mm, align=center, fill=oagray!40,
    inner sep=1pt},
  verdict/.style={align=left, font=\footnotesize},
  ]
  % The action.
  \node[align=center, font=\footnotesize\itshape] (act) {consequential\\action};
  \node[screen, right=10mm of act] (scr)
    {\textcolor{oagreen}{$\checkmark$}\,``Saved''\\\scriptsize green banner};
  \node[db, right=20mm of scr] (db) {system of\\record};

  \draw[-{Stealth[length=2mm]}, thick, oaink] (act) -- (scr);
  \draw[-{Stealth[length=2mm]}, thick, oaink] (act.east) to[out=-30,in=180] (db.west);

  % Weak vs strong oracle labels.
  \node[above=1mm of scr, font=\scriptsize\itshape, text=oared] {weak oracle (screen)};
  \node[above=1mm of db, font=\scriptsize\itshape, text=oagreen!55!black] {strong oracle (state)};

  % The disagreement.
  \node[verdict, below=7mm of scr, text width=78mm, xshift=18mm] (gap)
    {\textcolor{oared}{Same green banner} can follow: partial save $\cdot$
     duplicate $\cdot$ stale overwrite $\cdot$ silent reject.
     \\[1mm]
     \textcolor{oagreen!55!black}{Effect verifier} snapshots state before/after
     $\Rightarrow$ \textbf{confirmed} / \textbf{refuted} / \textbf{indeterminate};
     a refuted effect \emph{halts}.};
  \draw[-{Stealth[length=1.6mm]}, oared, dashed] (scr.south) -- (gap.north -| scr.south);
  \draw[-{Stealth[length=1.6mm]}, oagreen!55!black, dashed] (db.south) to[out=-90,in=0] (gap.east);
\end{tikzpicture}
\caption{Effect verification separates the weak screen oracle from the strong
system-of-record oracle. A rendered success is evidence about pixels, not about
a persisted write; only an independent before/after snapshot of the application's
state can refute a silent wrong action. In the fault study this reduced silent
acceptance from 50 of 90 runs to 0 of 90.}
\label{fig:oracle}
\end{figure}

\subsection{Policy and refusal}
The linter exposes unarmed clicks, vacuous postconditions, and risk-classifier
gaps. Certification evaluates a named policy before deployment. A strict run
gate can require certification, identity and effect coverage, encryption, and
manifest integrity. Passing a policy means only that its assertions hold; it is
not a general safety proof.

\subsection{Sanitized artifact derivatives}
Compilation is not de-identification. When an artifact crosses a boundary, the
launch protocol inventories the source, transforms a copy with type-specific
handlers, rescans the derivative, records unresolved findings and tool versions,
and binds local review to the derivative hash. Unknown or unresolved content is
refused. Sanitized authoring evidence is distinct from runtime observations:
opening a live patient record can reintroduce protected information, which must
remain inside the declared trusted execution boundary.
