\section{Evaluation Methodology}
\label{sec:method}

We separate comparative task performance, breadth, transactional safety, and
identity availability. A single aggregate success rate would hide materially
different errors.

\paragraph{Comparative conditions.}
Both repeated-workflow comparisons were generated on 2026-07-08 on the recorded
platform \texttt{macOS-15.7.3-arm64-arm-64bit}. The compiled and agent arms
drove the same vision-only Playwright backend: screenshots in and pixel clicks,
typed text, key presses, and scrolling out, with no DOM selectors at runtime.
The agent condition used \texttt{claude-sonnet-5} with the
\texttt{computer\_20251124} tool and
\texttt{computer-use-2025-11-24} beta header. Agent costs are calculated from
recorded token usage and the list-price schedule stored in each raw result; the
compiled condition recorded no model calls. The agent is a single-attempt
observe-act loop given the task goal and the same settled screenshots the
replayer uses, without a system prompt, planning or reflection scaffolding, or
whole-task retry, and bounded by per-run action and cost caps. It is a fair
same-backend reference rather than a maximally scaffolded agent, so the latency
and cost gaps should be read as lower bounds on what scaffolding could close,
not as a reliability comparison.

\paragraph{Outcome taxonomy.}
Success means the external task oracle is satisfied. A safe halt stops without
the intended effect. Over-halt is a safe halt when the intended correct action
was in fact available. Silent incorrect success means the runtime reports
success while the external oracle disagrees. False abort means the intended
effect occurred but the runtime reports failure. We report run counts rather
than confidence intervals for small bounded studies.

\paragraph{OpenEMR comparison.}
The task logs into the shared public OpenEMR demonstration site, searches a
patient, opens Patient Messages, adds a parameterized note, and saves. The
compiled arm has 20 runs and the computer-use-agent arm 10. Both use the same
task-level success check. The shared public instance is mutable and the study is
not fully reproducible in CI; it is field evidence, not a production study.

\paragraph{MockMed comparison.}
The bundled browser fixture performs a triage workflow. The compiled arm has
100 runs and the agent arm 20. This is the CI-reproducible anchor, but the
application is purpose-built and is not evidence about hostile enterprise UIs.

\paragraph{Public-web reliability corpus.}
Twenty-nine public, unauthenticated applications are recorded, compiled, and
replayed once on an unchanged UI using an external goal oracle. This study is
descriptive: one run per application does not meet the repeated-trial standard
for a capability claim, and the selection excludes SSO-walled and desktop
targets.

\paragraph{Transactional fault model.}
Seven persistence faults and controls are injected at a real HTTP persistence
boundary behind the MockMed UI, with 10 repeats per class. Database state, not
the UI report, is the oracle. This per-class study measures what screen-only
verification silently mishandles; it is separate from the end-to-end study below.

\paragraph{End-to-end effect verification.}
The headline safety result is measured through the actual governed replay path,
not an in-process stub. Ten persistence scenarios, seven fault classes plus an
accepted-write control and two delivery controls, are each run nine times per
arm. Every write is issued by the replayer's actuator as an HTTP POST to an
on-disk SQLite system of record, and each run is judged by three genuinely
distinct paths: the write itself; an effect verifier that reads the record back
over a different endpoint and connection than the write; and an independent
ground truth that opens the SQLite file read-only over every table and classifies
the before/after state without ever seeing the write's success flag. Because the
verifier and the ground truth share neither code nor state, the effect
verifier's zero is earned rather than definitional. We evaluate three arms:
screen-only banner verification, an out-of-band effect oracle whose read path
covers the record surface, and a complete effect oracle whose read path covers
every mutable surface the action can touch. This supersedes an earlier in-process
study in which the effect verifier and the ground truth read the same object,
which made its zero circular by construction. Zero model calls, localhost only.

\paragraph{Second domain: lending.}
To test generalization beyond healthcare, the same effect-verification protocol
is replicated on MockLoan, a loan-disbursement authorization workflow with its
own ledger. Eleven tasks spanning all seven divergence categories plus clean and
idempotent controls are each run three times per arm, judged by an independent
REST verifier that reads the ledger at an endpoint the application never calls.
One task seeds a same-name decoy loan to exercise the identity gate on the
money-movement step. Zero model calls.

\paragraph{Released benchmark.}
The metric and fault taxonomy are packaged as EffectBench, a standalone,
versioned, and independently runnable benchmark (specification version 1.0.0), so
that the silent-wrong-effect rate can be measured against any GUI or API
automation, not only ours.

\paragraph{Identity ladder.}
Adversarial same-name/same-date-of-birth homonyms differ by confusable glyphs
such as O/0 and l/1. The actual replayer stack is evaluated with structured and
pure-pixel configurations. False accept and over-halt are both reported because
a system can obtain zero wrong clicks by refusing every correct one.

\paragraph{Desktop and remote-substrate qualification.}
These mechanism studies are task-bound rather than comparative. Windows UIA is
evaluated on one in-tree WinForms Patient Notes workflow in a Windows 11 ARM VM.
Each counted trial starts from a clean snapshot, writes a trial-unique note, and
uses an independent SQLite query as the effect oracle; stale and ambiguous
candidate controls must leave every row unchanged. Native macOS is evaluated on
one macOS 15.7.3 arm64 host using TextEdit, exact file-byte readback, and a
two-window ambiguity control. The preserved batch report classifies graceful-
close cleanup warnings as failure; a separate hash-bound offline adjudication
accepts only the three action effects and ambiguity refusal after verifying the
exact harness processes and temporary root were absent. RDP is evaluated over
Aardwolf 0.2.14 into one 1280x800 Parallels Windows 11 snapshot: input through
the Windows Run dialog creates a trial-unique file whose exact contents are read
back through guest tools independently of the RDP channel. Each promoted task
has three counted trials, records model calls, silent incorrect success and
over-halt, and retains environment-cleanup evidence.
