\section{Reproducibility and Artifacts}

The source repository and reviewed public evaluation artifacts are published
together \cite{openadaptartifact} at \artifacturl{}. The paper package includes
a script that reads released benchmark JSON and fails if a headline constant
drifts. Depending on the study, that released source is either a raw result or a
bounded aggregate summary. Task definitions, benchmark prose, failure
taxonomies, and selected run reports are versioned where their release boundary
permits it. Grown identity and reliability corpora, deployment-derived tuning,
target-specific recipes, and raw private rows are deliberately excluded. The
machine-check therefore establishes consistency between the paper and released
evidence; it neither claims that every underlying private row is public nor
validates the design of the underlying measurement.

Evidence carries one of four labels:
\begin{itemize}
  \item \textbf{CI-reproducible}: fixture and environment can run in automated
        CI, subject to browser and dependency availability.
  \item \textbf{Field}: real third-party target, but shared mutable state or
        credentials prevent full CI reproduction.
  \item \textbf{Fixture/fault injection}: deterministic mechanism study, not a
        population estimate.
  \item \textbf{Descriptive}: useful for discovering failures, but missing the
        repeated trials needed for comparative inference.
\end{itemize}

Before submission, the release candidate must re-run promoted experiments,
archive its exact commit and environment, record task-level oracles, and check
that no application artifact contains sensitive or prohibited data. The
repository's \texttt{paper/ARTIFACT\_CHECKLIST.md} is the submission gate.

\paragraph{Data availability and ethics.}
Every released system-of-record fixture is synthetic: the MockMed and MockLoan
fixtures, their injected-fault suites, and the benchmark task packs contain no
real patient, borrower, or customer data. The public-web study observed public,
unauthenticated applications, but only its bounded aggregate is released; its
raw per-target rows and curated target recipes remain private. This paper
release contains no customer-derived hardening corpus. Such grown corpora and
raw private evaluation rows remain outside the public repository; when an
approved sanitized derivative crosses a trust boundary, the pipeline applies
type-specific transformations, rescans the result, and binds human approval to
its exact hash as described in
Section~\ref{sec:governance}. Human review reduces but does not prove
de-identification, so sanitization remains a defense-in-depth control rather
than a guarantee. Repository-only third-party reference environments are also
excluded from the permissively licensed wheel and source distribution; the
release gate inspects the built archives rather than inferring their contents
from the source tree.
