\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,shapes.geometric,fit,backgrounds,calc,decorations.pathreplacing}
\usepackage{pgfplots}
\pgfplotsset{compat=1.13}

% Shared figure palette.
\definecolor{oaink}{HTML}{1F2933}
\definecolor{oablue}{HTML}{2563EB}
\definecolor{oagreen}{HTML}{15803D}
\definecolor{oared}{HTML}{B91C1C}
\definecolor{oaamber}{HTML}{B45309}
\definecolor{oagray}{HTML}{E5E7EB}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=blue,
  urlcolor=blue,
  pdftitle={Compile Once, Govern Every Repair: Deterministic Replay for Repeated GUI Work},
  pdfsubject={A reproducible technical-report draft for OpenAdapt}
}
\newcommand{\system}{OpenAdapt}
\newcommand{\artifacturl}{\url{https://github.com/OpenAdaptAI/openadapt-flow}}

\title{Compile Once, Govern Every Repair:\\Deterministic Replay for Repeated GUI Work}
\author{Richard Abrich\\OpenAdapt.AI (MLDSAI Inc.)\\\texttt{richard@openadapt.ai}}
\date{Draft of July 2026}

\begin{document}
\maketitle

\begin{abstract}
Reasoning through a known workflow on every run is wasteful and unsafe. It adds
latency, inference cost, and run-to-run variation, and---because a rendered
success banner is not a persisted write---it can report success after a partial,
duplicate, stale, or rejected business effect. We present \system{}, a
demonstration compiler that converts one recorded GUI trace into a deterministic
program. Healthy replay follows structured elements and local visual evidence
with \emph{no model calls}; when evidence changes, a resolution ladder
re-resolves the target and records the change as a reviewable \emph{repair}; and
the runtime \emph{refuses} to continue when configured identity, postcondition,
effect, or policy checks cannot verify an action. The central safety mechanism is
system-of-record effect verification: a consequential step declares a typed
effect and checks it against the application's own state, not the screen.
Measured end to end through the live replayer and judged by an independent
system-of-record ground truth, screen-only verification silently accepted 54 of
90 injected-fault runs; an out-of-band effect check cut that to 9 of 90 and a
complete system-of-record read path to 0 of 90, the residual nine being one
collateral-write class outside the out-of-band oracle's read coverage. The
finding replicates on a second, non-healthcare lending domain: 21 of 33 silent
wrong effects under screen-only verification, 0 of 33 under effect verification.
Bounded experiments span a public OpenEMR
workflow, a CI-reproducible fixture, a 29-application public-web corpus,
transactional fault injection, adversarial identity tests, and scoped
qualifications of native Windows UIA, macOS, and real-network RDP mechanisms.
Across these tasks compiled replay reduced latency and per-run inference while
preserving independently checked effects and ambiguity refusal. Every headline
number in this paper is bound by a released machine-check to its raw artifact;
the release publishes the implementation, reviewed artifacts, and failure
taxonomy.
\end{abstract}

\input{sections/01_introduction}
\input{sections/01_related_work}
\input{sections/02_system}
\input{sections/03_governance}
\input{sections/04_methodology}
\input{sections/05_results}
\input{sections/06_limitations}
\input{sections/07_reproducibility}
\input{sections/08_conclusion}

\bibliographystyle{plain}
\bibliography{references}
\end{document}
